### 2020-08-20 text/article_vedrovice.Rmd
cont. art. text - spatial patterning

Given statistics for independent random process are estimated using Monte Carlo
simulation with 99 iterations.
The most extreme results of simulations are discarded as outliers and the
second most extreme results are used to construct envelopes encompassing the
expected range for a given function [@baddeley2016 384-403].

@@

The quadrat test suggests both random and clustered processes with p values
under the 0.05 alpha value but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Fig. \@ref(fig:spat-funs) shows that the curves of $G$ and $F$ functions do not
deviate too much from the curves expected under independent random process nor
do they traverse the boundaries of the envelopes.
This tells us that both the cummulative nearest-neighbor distance (*G*) and
distance to any random location in the space of the window (*F*) are well
approximated by random processes and do not show evidence for spatial clustering.
On the other hand, both *K* and *T* curves have a steep rise and exceed the
envelope boundaries.
These functions thus show evidence for clustering process in the observed point
pattern.

>>>>

Given statistics for independent random process are estimated using Monte Carlo
simulation with 999 iterations.
Ten most extreme realizations of the simulation are discarded as outliers to
get a significance level $\alpha = 0.01$ [@baddeley2016 384-403].

@@

The quadrat test shows evidence for both random and clustered processes taking place with p values
under the 0.05 alpha value but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Inconclusive result of the quadrat test is not suprising because as was noted
earlier, we do not expect *first-order* properties to play major roles in
placing of the individual burials.
Results of functions based on nearest neighbors are reported in Fig. \@ref(fig:spat-funs).
*G* function does deviate from the curve expected under independent random process
but does not traverse the boundaries of the envelope.
Thus the cummulative nearest-neighbor distance (*G*) is well approximated by
random process and does not show evidence for spatial clustering nor repelent functions
taking place.
The empty-space function *F* deviates slightly from the area defined by
the envelope at around 6 meters.
This means that the number of nearest neighbours to any empty-space point
at this and larger distances is lower than it would be expected under complete spatial randomness.
On the contrary to only slight deviations in case of *G* and *F*
the *K* and *T* functions show clear deviations from the expected curve and exceed
the boundaries of given envelopes at the distance of around 5 to 6 meters in
case of *K* and slightly more in case of *T* function.
This means that numbers of nearest neighbors at this and larger distances is
significantly higher than we would exect under complete spatial randomness.
The overall point pattern thus shows clear signs of clustering process taking place.

This means that the locations of new burials were not chosen at random
but were affected by the positions of previous burials.
The pattern begins to cluster at the distance of around 5 to 6 meters, meaning
that beginning at this distance, the probability of finding a neighboring burial is
much higher than it would be if the burials were placed randomly.
The community burying their dead at the cemetery of Vedrovice -- Široká u lesa
did so in an organized manner taking into account previous burials at the same place,
meaning their positions were known, probably marked on the surface somehow.
This would imply either non-perishable marks on the surface of the cemetery or their
maintenance from time to time.

Spatial pattern was also explored for differently sexed bodies.
